High resolution, annual maps of the characteristics of smallholder-dominated croplands at national scales

### 2021-03-09 inst/manuscript/manuscript.Rmd
Edited ms and si: first upload to EarthArXiv. Edited scripts for data_for_instances.R, added new script to data-raw (aois_tiles.R) to show creation of degree and tiling grids

- name: L.D. Estes
  affilnum: '1'
  email: lestes@clarku.edu

- name: S. Ye
  affilnum: '1,2'

- name: L. Song
  affilnum: '1'

- name: B. Luo
  affilnum: '1,3'

@@

- name: Z. Meng
  affilnum: '1'

- name: Q. Zhang
  affilnum: '1'

- name: D. McRitchie
  affilnum: '4'

- name: S. Debats
  affilnum: '4'

- name: J. Muhando
  affilnum: '5'

- name: A. Hokar
  affilnum: '5'

- name: C. Juma 
  affilnum: '5'

- name: J. Makuru
  affilnum: '5'

- name: F. Matasa
  affilnum: '5'

- name: B. Mbatia
  affilnum: '5'

- name: J. Mbaye
  affilnum: '5'

- name: I. Muasa
  affilnum: '5'

- name: J. Mucha
  affilnum: '5'

- name: A. Mugami
  affilnum: '5'

- name: P. Oduor
  affilnum: '5'

- name: J. Ochieng
  affilnum: '5'

- name: F. Wambua
  affilnum: '5'

- name: T. Wanjiku
  affilnum: '5'

- name: J. Wanyoike 
  affilnum: '5'

- name: R. Avery
  affilnum: '6'

- name: K. Caylor
  affilnum: '6,7,8'

@@

Which is a weighted mean for each pixel for each band *B* for the particular compositing period. The composited tiles were then added to the S3 store (Figure \ref{fig:systemoverview} 1), where they are stored as cloud-optimized geotiffs, and a "slippy map\footnote{https://wiki.openstreetmap.org/wiki/Slippy_Map}" rendering is created for each composite using Raster Foundry$\footnote{https://github.com/raster-foundry/raster-foundry}$. The web-rendered imagery is presented within the training data platform (next section).

### Labelling platform
Training and reference data are collected by a custom labelling platform, which was originally designed for AWS's Mechanical Turk job marketplace [@Estesplatformcrowdsourcingcreation2016]. The basic structure of the system remains the same, but we converted it into a standalone platform that allows us to enroll and pay people directly for their labelling, and is designed to control and supervise the machine learning process. The platform runs on a Linux virtual machine hosted on an AWS EC2 instance and is comprised of a database (PostGIS/Postgres), a mapping interface (OpenLayers 3), an image server (RasterFoundry), and a set of utilities for managing, assessing, and converting digitization work into rasterized labels for training a machine learning algorithm. Each instance of the platform focuses on a specific AOI (Figure \ref{fig:aois}A)

@@

```{r labeller, echo = FALSE, out.width="100%", fig.cap="An overview of the labelling platform's interface", fig.align='center'}
knitr::include_graphics('figures/figure3.png')
```

@@

\begin{center}Table 1. List of features.\end{center}

Feature             Window Size  N Features
------------------- ------------ ------------
RGB-NIR                      1X1            8 
Mean                       11X11            8
Standard deviation           5X5            8
------------------ ------------- ------------

Feature extraction and the conversion of image features is handled by a combination of `GeoTrellis`\footnote{https://github.com/locationtech/geotrellis}, `rasterio`\footnote{https://rasterio.readthedocs.io/en/latest/}, and `RasterFrames`\footnote{https://www.cogeo.org/}. These collectively extract subsets of imagery from the PlanetScope temporal composites, derive the features, and convert these into `RasterFrames`, a data structure designed for handling raster data within Apache Spark. Features are extracted on the fly for each cell in the training and validation sets, a functionality enabled by storing the image composites as Cloud-optimized Geotiffs\footnote{https://www.cogeo.org/} (COGs), which allows rapid reading of image subsets.

#### Classification

Once the features from the training sites are extracted into `RasterFrames`, these are combined with their corresponding labels and passed to the machine learning classifier, a `SparkMLlib` implementation of Random Forests [@BreimanRandomForests2001]. For this study, the model was trained with a balanced sample and a tree depth of 15 and total tree number of 60, which initial testing showed to provide a reasonable balance between computational time and performance.

@@

To develop the maps, we first generated an image catalog for the `r nt` tiles covering Ghana. This entailed processing all PlanetScope imagery intersecting these tiles between May-September, 2018 (the growing season) and December, 2018 to February, 2019 (the subsequent dry season). The longer period was necessary for the growing season because of the frequent cloud cover, which severely limits the number of clear scenes for any tile. For the cloudiest regions (AOIs 10, 11, 13, 14, 16) we started the dry season window in November.

An assessment with two observers (see SI for observer details) found that average quality per growing season composite tile was `r tqmu["Growing"]`, with `r tqct["Growing"]` percent having scores $\geq$ 0.85, while the average quality of dry season composites was `r tqmu["Dry"]` (`r tqct["Dry"]` percent $\geq$ 0.85). Composite quality in both seasons was highest in the northern half of the country and lowest in the southwest (Figure \ref{fig:imqual}A), where the substantially greater cloud cover resulted in a much lower density of available PlanetScope imagery for each time period.

@@

After training of models with the initial randomly selected label sets, the active learning process was run for 3 iterations for 12 of the 16 AOIs, resulting in `r typicaln` labels per AOI. AOIs 10 and 14 stopped after one and two iterations, respectively, as they started with high initial validation accuracies (>83%) and showed little subsequent improvement. The models for these two AOIs were thus trained with `r aoi10n` -  `r aoi14n` samples. AOI 15 was run for 4 iterations (`r aoi15n` samples), while AOI 3 underwent a second active learning cycle because the model produced during the first cycle was inaccurate (see SI). In this second run, `r aoi3init` initial training sites randomly selected within the AOI were used (Figure S3A), followed by 2 subsequent active learning iterations, resulting in a training sample of `r aoi3n`. Labels collected during the active learning iterations showed distinct patterns in several AOIs, which often fell along ecotones, such as the boundaries between agroecozones (see Figure S3A). The total number of unique training and validation sites across the country were `r ttrain` and `r tval`, respectively.

@@

The distribution of training and validation sample collection effort was divided across `r nrow(label_summary$score_stats)` labellers, with a core group of `r label_summary$nlabels %>% filter(kml_type == "Training" & n > 1000) %>% nrow()` who mapped more than 1,000 sites each. As each training site was mapped by 4 separate labellers, `r tsitesc$tot[1]` sets of vector labels were made. Each labeller digitized an average of `r tsitesc$mu[1]` (see Figure S4A for more details on labelling effort). Labeller accuracy was scored `r tsitesc$tot[2]` times against `r length(unique(label_summary$scores$name))` unique training reference sites (Figure S3A), with each labeller assessed an average of `r tsites$mu[2]` times at a rate of 1 training reference site for every `r round(1 / (tsites$mu[2] / tsites$mu[1]), 2)` training site mapped. The mean of each labeller's average accuracy score was `r round(mean(label_summary$score_stats$Mean), 2)` (range `r round(min(label_summary$score_stats$Mean), 2)` to `r round(max(label_summary$score_stats$Mean), 2)`; see Figure S4B for detailed score distributions).

@@

After each site was mapped by four labellers, consensus labels were generated. The Bayesian Risk (see SI) of each training and validation label was calculated as an additional measure of label quality. The average risk was `r murisk$Train` for training labels and `r murisk$Validate` for validation labels. Risk was highest in the northen AOIs (AOIs 1-6; Figures S5-6), falling between `r cl1$Risk[1]` for training and `r cl1$Risk[2]` for validation labels (Figures S5-6), and lowest in the southwestern AOIs (AOIs 10, 11, 13, 14, 16; training risk = `r cl3$Risk[1]`; validation risk = `r cl3$Risk[2]`). Label risk in the central-southeastern AOIs (AOIs 7-9, 12, 15) was slightly lower (training = `r cl2$Risk[1]`; validation = `r cl2$Risk[2]`) than in the north. Labeller experience also appeared to reduce risk, which we observed during a relabelling of the 500 initial random site in this cluster (see SI); the mean risk of the updated labels was `r cl2brisk`, compared to `r cl2arisk` for original labels.

@@

```{r alperformance, echo = FALSE, out.width="100%", fig.cap="Scores for overall Accuracy, area under the curve of the Receiver Operating Characteristic, and the F1 scores for the Random Forests model results after each iteration of the active learning loop for each AOI (gray lines), as well as the mean score per iteration across all AOIs (black lines).", fig.align='center', message=FALSE}
knitr::include_graphics('figures/figure5.png')
```

@@

The comparison of active versus randomized training sample collection (in AOIs 1, 8, and 15) showed that the former approach outperformed the latter. After three iterations, the accuracy, AUC, and F1 scores resulting from active learning were respectively `r u(ixl$III["Accuracy"])`, `r u(ixl$III["AUC"])`, and `r u(ixl$III["F1"])` percent higher than the scores from a randomly trained model (Figure S7). However, there was more variability in earlier iterations, with average score differences of `r u(ixl$I["Accuracy"])` (accuracy), `r u(ixl$I["AUC"])` (AUC), and `r u(ixl$I["F1"])` percent (F1) after iteration 1, and `r u(ixl$II["Accuracy"])` (accuracy), `r u(ixl$II["AUC"])` (AUC), and `r u(ixl$II["F1"])` (F1) percent after iteration 2. The negative results for accuracy was caused by results at AOI 15, where active learning accuracy was `r abs(u(i1acc["15"]))` percent lower than random training after iteration 1 (see Figure \ref{fig:alperformance}). In comparison, iteration 1 active learning accuracies were `r u(i1acc["1"])` and `r u(i1acc["8"])` percent higher than random training for AOIs 1 and 8, respectively. Accuracy under active learning for AOI 15 exceeded randomized training after 4 iterations.

@@

We found that the overall accuracy of the pixel-wise classifications was 88% against this map reference sample (Table \ref{tab:mapaccuracy}). Confining the map reference sample to four distinct zones (Figure S9A) shows that overall accuracy ranged from 83.3% in Zone 1 (AOIs 1-3) to 93.6% in Zone 3 (AOIs 10, 11, 13, 15, and 16). The Producer's accuracy of the cropland class was 61.7% across Ghana, ranging from 45.6% in Zone 3 to 67.9% in Zone 1, while the User's accuracy for was 67.3% overall, ranging from 59.8% in Zone 4 to 71.2% in Zone 1. Both measures of accuracy were substantially higher for the non-cropland class across all zones, typically exceeding 90%. The lowest accuracies for the non-cropland class was in Zone 1 (Producer's = 89.3%; User's = 87.7%).

@@

```
The comparisons of digitized versus segmented field boundaries showed that the mean field size across all validation sites averaged `r round(segha, 2)` ha (Median = `r round(seghamed, 2)`; StDev = `r round(seghasd, 2)`), which was `r round((segha - valha) / valha * 100) / 100` times larger than the `r round(valha, 2)` ha (Median = `r round(valhamed, 2)`; StDev = `r round(valhasd, 2)`) mean area of labeller-digitized polygons. This discrepancy was primarily caused by results in four AOIs (2, 3, 7, and 15; Figure S10), where segments averaged between `r round(lrgst_seg$rng[1], 2)` and `r round(lrgst_seg$rng[2], 2)` ha, compared to `r round(lrgst_val$rng[1], 2)` - `r round(lrgst_val$rng[2], 2)` ha for the corresponding hand-digitized polygons. The number of segmented fields per validation site averaged `r segn` (median = `r segnmed`; StDev = `r segnsd`) compared to `r valn` (median = `r valnmed`; StDev = `r valnsd`) for digitized polygons (Figure S11).

@@

The vectorized map provides additional information on how the characteristics of croplands can vary geographically, ranging from narrow, strip-like fields in parts of AOI 15 (Figure \ref{fig:mainmap}'s lower right inset) to more densely packed, less distinctly shaped fields in AOI 5 (upper right inset in Figure \ref{fig:mainmap}). To explore how field characteristics varied geographically, we mapped the average field size and total number of fields within each 0.05 degree tile (Figure S12). These patterns generally correspond to those seen in the cropland density map (Figure \ref{fig:mainmap}), with larger sizes and field counts generally occurring in areas of higher field density, although the biases inherent in both measures (Figures S10-11) complicate the interpretation of those variations. However, we can use the estimated biases to develop adjusted estimates of field sizes and counts for each AOI, and for Ghana overall (Table \ref{tab:sizentab}). These adjusted estimates show that the typical field size in Ghana is 1.73 ha, ranging from 0.96 in AOI 4 to 2.82 ha in AOI 4, with fields in the forest zone AOIs (10, 11, 13, 14, 16) generally smaller than those in the northern half of the country (Table \ref{tab:sizentab}). The total number of fields was estimated to be 1,662,281 overall, or 205 fields per tile on average, ranging from 108/tile in AOI 4 to 399/tile in AOI 6.

@@

In addition to the more detailed update of cropland extent, our maps also provide new information on the size and number of fields in Ghana (Figures \ref{fig:mainmap}, S11-12). Previous work to estimate such agricultural characteristics have often focused on farm, rather than field, size using census data [@jayneAfricaChangingFarm2016; @LowderNumberSizeDistribution2016; @SambergSubnationaldistributionaverage2016; @VonBraunSmallscalefarmersliberalized2004]. Efforts to map field boundaries in smallholder-dominated agricultural systems have either used *in situ* data collection [@CarlettoFactartifactimpact2013; @CarlettoGuesstimatesGPStimatesLand2015] or remote sensing studies over relatively small [e.g. @forkuorIntegrationOpticalSynthetic2014; @perselloDelineationAgriculturalFields2019] or discontiguous [@Estesplatformcrowdsourcingcreation2016] areas. The most extensive studies to date used crowdsourced volunteers to classify fields into broad size classes, based on their interpretations of imagery sampled from high resolution virtual globes [@lesivEstimatingGlobalDistribution2019; @FritzMappingglobalcropland2015]. Those efforts included country-specific results for Ghana (n = 263), which can be converted into an average field size estimate of `r round(lesiv_est, 2)` ha$\footnote{Obtained by calculating the weighted mean from the count of the five size classes and the mean of the hectare range provided for the four smallest size classes, and the lower bound of the size range provided the largest size class. Data sourced from Table S3 in Lesiv et al. 2019.}$. This estimate exceeds our Ghana-wide average segment size (3.92 ha; Table \ref{tab:sizentab}), but is closer to the mean (`r aoi_nofor` ha) within AOIs 1-9, 12, and 15, which is where most of the crowdsourced sample appears to have been collected. However, our bias-corrected estimates of 1.73 (Ghana-wide) and `r aoi_nofor_adj` (AOIs 1-9, 12, and 15) ha were much smaller.

@@

The patterns of cropland-class accuracies varied by zone. These zones largely align, albeit with some discrepancies, with the country's agroecozones, thus the accuracy patterns may be partially attributed to some regions being harder to map than others. Producer's accuracies for both maps were highest in the two northern zones (1 and 2), which are primarily savannas (Figure S9), and lowest in zones 3 and 4, which are comprised of forest or coastal savannas. User's accuracies followed a similar pattern, with the exception of Zone 3, which had the highest User's accuracy, albeit from a very small sample. Aligning the reference samples more precisely with agroecozone boundaries (Figure S9B) provides further insight into error patterns (Table S4). Coastal savannas in the southeast had the highest Producer's accuracy but lowest User's accuracy for the per-pixel map, presumably because this region's numerous areas of high density cropland, combined with low woody cover in surrounding uncultivated areas, helped to promote commission error. Maps in the two northern savanna agroecozones had the best balance between omission and commission error, and had the highest overall User's accuracy. The transitional (between forest and savanna) agroecozone had a very low Producer's accuracy (21%), which likely reflects the fact that it was divided between several AOIs for mapping (Figure S3), within which it typically covered a smaller share of area relative to the other agroecozones. This likely caused insufficient representation of this AEZ in training samples, particularly in AOIs 10 and 11 (Figure S3B).

@@

The inherent difficulty of the labelling task was another major limiting factor. Our system was designed to minimize the error inherent in labelling, but determining croplands from non-croplands in these agricultural systems can be a difficult task. Labellers have to evaluate multiple image sources and to rely heavily on judgement, which inevitably leads to errors. Interpretation is particularly hard where the background savanna vegetation and croplands have similar reflectance during the dry season, which is a particular problem in AOIs 2 and 3. Smaller field sizes also complicate labelling, as these become increasingly indistinct in the ~4 m PlanetScope composites. The difficulty of labelling is reflected in the magnitude of the Bayesian Risk metrics (Figure S5), and by the average score achieved by each labeller against our training reference dataset  (71%; Figure S4B). Although prior work [@mellorExploringIssuesTraining2015; @rodriguez-galianoAssessmentEffectivenessRandom2012] found that Random Forests are robust to label errors, we found that they have substantial impact (Figure \ref{fig:trainingimpact}), which suggest that simply improving label quality may be one of the single most important investments towards improved model accuracy.

@@

## Error mitigation features
Despite the error sources mentioned above, several features of the system proved effective in mitigating error, leading to a higher overall accuracy than would have otherwise been possible. Label accuracy assessment and consensus labelling appeared to be the most effective error mitigation tools. Label accuracy measures allowed us to quantify the substantial impact of label error on model performance (Figure \ref{fig:trainingimpact}), while consensus labels substantially reduced individual labelling errors, resulting in maps that were more accurate than they would have been had we used individually generated labels. Labeller-specific accuracy measures also helped to improve the quality of the consensus labels, by placing higher weight on labels more likely to be accurate during the merging process, rather than giving equal weight to potentially less accurate labels. The ability to select the most accurate individual labels for a site also allowed us to develop independent estimates of field size to which measures of confidence can be attached (Figure S4B), which we were in turn able to correct estimates of field sizes and numbers (Table \ref{tab:sizentab}).

>>>>

- name: Lyndon D. Estes
  affilnum: '1'
  email: lestes@clarku.edu

- name: Su Ye
  affilnum: '1,2'

- name: Lei Song
  affilnum: '1'

- name: Boka Luo
  affilnum: '1,3'

@@

- name: Zhenhua Meng
  affilnum: '1'

- name: Qi Zhang
  affilnum: '1'

- name: Dennis McRitchie
  affilnum: '4'

- name: Stephanie R. Debats
  affilnum: '4'

- name: Justus Muhando
  affilnum: '5'

- name: Angeline H. Amukoa
  affilnum: '5'

- name: Brian W. Kaloo
  affilnum: '5'

- name: Jackson Makuru
  affilnum: '5'

- name: Ben K. Mbatia
  affilnum: '5'

- name: Isaac M. Muasa
  affilnum: '5'

- name: Julius Mucha
  affilnum: '5'

- name: Adelide M. Mugami
  affilnum: '5'

- name: Judith M. Mugami
  affilnum: '5'

- name: Francis W. Muinde
  affilnum: '5'

- name: Fredrick M. Mwawaza
  affilnum: '5'

- name: Jeff Ochieng
  affilnum: '5'

- name: Charles J. Oduol
  affilnum: '5'

- name: Purent Oduor
  affilnum: '5'

- name: Thuo Wanjiku
  affilnum: '5'

- name: Joseph G. Wanyoike 
  affilnum: '5'

- name: Ryan B. Avery
  affilnum: '6'

- name: Kelly K. Caylor
  affilnum: '6,7,8'

@@

Which is a weighted mean for each pixel for each band *B* for the particular compositing period. The composited tiles were then added to the S3 store (Figure \ref{fig:systemoverview}), where they are stored as cloud-optimized geotiffs, and a "slippy map\footnote{https://wiki.openstreetmap.org/wiki/Slippy\_Map}" rendering is created for each composite using Raster Foundry [@azaveaRasterFoundry2020]. The web-rendered imagery is presented within the training data platform (next section).

### Labelling platform
Training and reference data are collected by a custom labelling platform, which was originally designed for AWS's Mechanical Turk job marketplace [@Estesplatformcrowdsourcingcreation2016]. The basic structure of the system remains the same, but we converted it into a standalone platform that allows us to enroll and pay people directly for their labelling, and is designed to control and supervise the machine learning process. The platform runs on a Linux virtual machine hosted on an AWS EC2 instance and is comprised of a database (PostGIS/Postgres), a mapping interface (OpenLayers 3), an image server (Raster Foundry), and a set of utilities for managing, assessing, and converting digitization work into rasterized labels for training a machine learning algorithm. Each instance of the platform focuses on a specific AOI (Figure \ref{fig:aois}A)

@@

```{r labeller, echo = FALSE, out.width="95%", fig.cap="An overview of the labelling platform's interface", fig.align='center'}
knitr::include_graphics('figures/figure3.png')
```

@@

\begin{center}Table 1. List of image features.\end{center}
Feature             Window Size  N Features
------------------- ------------ ------------
RGB-NIR                      1X1            8 
Mean                       11X11            8
Standard deviation           5X5            8
------------------ ------------- ------------
Feature extraction and the conversion of image features is handled by a combination of `GeoTrellis`\footnote{https://github.com/locationtech/geotrellis}, `rasterio`\footnote{https://rasterio.readthedocs.io/en/latest/}, and `RasterFrames`\footnote{https://rasterframes.io/}. These collectively extract subsets of imagery from the PlanetScope temporal composites, derive the features, and convert these into Apache Spark DataFrames. Features are extracted on the fly for each cell in the training and validation sets, a functionality enabled by storing the image composites as Cloud-optimized Geotiffs\footnote{https://www.cogeo.org/} (COGs).

#### Classification
Once the features from the training sites are extracted into `RasterFrames`, these are combined with their corresponding labels and passed to the machine learning classifier, a `SparkMLlib` implementation of Random Forests [@BreimanRandomForests2001]. For this study, the model was trained with a balanced sample and a tree depth of 15 and total tree number of 60, which initial testing showed to provide a reasonable balance between computational time and performance.

@@

To develop the maps, we first generated an image catalog for the `r nt` tiles covering Ghana. This entailed processing all PlanetScope imagery intersecting these tiles between May-September, 2018 (the growing season) and December, 2018 to February, 2019 (the subsequent dry season). The longer period was necessary for the growing season because of the frequent cloud cover, which substantially limits the number of clear scenes for any tile (Figure S3). For the cloudiest regions (AOIs 10, 11, 13, 14, 16) we started the dry season window in November.

An assessment with two observers (see SI for observer details) found that average quality per growing season composite tile was `r tqmu["Growing"]`, with `r tqct["Growing"]` percent having scores $\geq$ 0.85, while the average quality of dry season composites was `r tqmu["Dry"]` (`r tqct["Dry"]` percent $\geq$ 0.85). Composite quality in both seasons was highest in the northern half of the country and lowest in the southwest (Figure \ref{fig:imqual}A), where the substantially greater cloud cover resulted in a much lower density of available PlanetScope imagery for each time period (Figure S3).

@@

After training of models with the initial randomly selected label sets, the active learning process was run for 3 iterations for 12 of the 16 AOIs, resulting in `r typicaln` labels per AOI. AOIs 10 and 14 stopped after one and two iterations, respectively, as they started with high initial validation accuracies (>83%) and showed little subsequent improvement. The models for these two AOIs were thus trained with `r aoi10n` -  `r aoi14n` samples. AOI 15 was run for 4 iterations (`r aoi15n` samples), while AOI 3 underwent a second active learning cycle because the model produced during the first cycle was inaccurate (see SI). In this second run, `r aoi3init` initial training sites randomly selected within the AOI were used (Figure S4A), followed by 2 subsequent active learning iterations, resulting in a training sample of `r aoi3n`. Labels collected during the active learning iterations showed distinct patterns in several AOIs, which often fell along ecotones, such as the boundaries between agroecozones (see Figure S4A). The total number of unique training and validation sites across the country were `r ttrain` and `r tval`, respectively.

@@

The distribution of training and validation sample collection effort was divided across `r nrow(label_summary$score_stats)` labellers, with a core group of `r label_summary$nlabels %>% filter(kml_type == "Training" & n > 1000) %>% nrow()` who mapped more than 1,000 sites each. As each training site was mapped by 4 separate labellers, `r tsitesc$tot[1]` sets of vector labels were made. Each labeller digitized an average of `r tsitesc$mu[1]` (see Figure S5A for more details on labelling effort). Labeller accuracy was scored `r tsitesc$tot[2]` times against `r length(unique(label_summary$scores$name))` unique training reference sites (Figure S4A), with each labeller assessed an average of `r tsites$mu[2]` times at a rate of 1 training reference site for every `r round(1 / (tsites$mu[2] / tsites$mu[1]), 2)` training site mapped. The mean of each labeller's average accuracy score was `r round(mean(label_summary$score_stats$Mean), 2)` (range `r round(min(label_summary$score_stats$Mean), 2)` to `r round(max(label_summary$score_stats$Mean), 2)`; see Figure S5B for detailed score distributions).

@@

After each site was mapped by four labellers, consensus labels were generated. The Bayesian Risk (see SI) of each training and validation label was calculated as an additional measure of label quality. The average risk was `r murisk$Train` for training labels and `r murisk$Validate` for validation labels. Risk was highest in the northen AOIs (AOIs 1-6; Figures S6-7), falling between `r cl1$Risk[1]` for training and `r cl1$Risk[2]` for validation labels (Figures S6-7), and lowest in the southwestern AOIs (AOIs 10, 11, 13, 14, 16; training risk = `r cl3$Risk[1]`; validation risk = `r cl3$Risk[2]`). Label risk in the central-southeastern AOIs (AOIs 7-9, 12, 15) was slightly lower (training = `r cl2$Risk[1]`; validation = `r cl2$Risk[2]`) than in the north. Labeller experience also appeared to reduce risk, which we observed during a relabelling of the 500 initial random site in this cluster (see SI); the mean risk of the updated labels was `r cl2brisk`, compared to `r cl2arisk` for original labels.

@@

```{r alperformance, echo = FALSE, out.width="100%", fig.cap="Scores for overall accuracy, area under the curve of the Receiver Operating Characteristic, and the F1 scores for the Random Forests model results after each iteration of the active learning loop for each AOI (gray lines), as well as the mean score per iteration across all AOIs (black lines).", fig.align='center', message=FALSE}
knitr::include_graphics('figures/figure5.png')
```

@@

The comparison of active versus randomized training sample collection (in AOIs 1, 8, and 15) showed that the former approach outperformed the latter. After three iterations, the accuracy, AUC, and F1 scores resulting from active learning were respectively `r u(ixl$III["Accuracy"])`, `r u(ixl$III["AUC"])`, and `r u(ixl$III["F1"])` percent higher than the scores from a randomly trained model (Figure S8). However, there was more variability in earlier iterations, with average score differences of `r u(ixl$I["Accuracy"])` (accuracy), `r u(ixl$I["AUC"])` (AUC), and `r u(ixl$I["F1"])` percent (F1) after iteration 1, and `r u(ixl$II["Accuracy"])` (accuracy), `r u(ixl$II["AUC"])` (AUC), and `r u(ixl$II["F1"])` (F1) percent after iteration 2. The negative results for accuracy was caused by results at AOI 15, where active learning accuracy was `r abs(u(i1acc["15"]))` percent lower than random training after iteration 1 (see Figure \ref{fig:alperformance}). In comparison, iteration 1 active learning accuracies were `r u(i1acc["1"])` and `r u(i1acc["8"])` percent higher than random training for AOIs 1 and 8, respectively. Accuracy under active learning for AOI 15 exceeded randomized training after 4 iterations.

@@

We found that the overall accuracy of the pixel-wise classifications was 88% against this map reference sample (Table \ref{tab:mapaccuracy}). Confining the map reference sample to four distinct zones (Figure S10A) shows that overall accuracy ranged from 83.3% in Zone 1 (AOIs 1-3) to 93.6% in Zone 3 (AOIs 10, 11, 13, 15, and 16). The Producer's accuracy of the cropland class was 61.7% across Ghana, ranging from 45.6% in Zone 3 to 67.9% in Zone 1, while the User's accuracy for was 67.3% overall, ranging from 59.8% in Zone 4 to 71.2% in Zone 1. Both measures of accuracy were substantially higher for the non-cropland class across all zones, typically exceeding 90%. The lowest accuracies for the non-cropland class was in Zone 1 (Producer's = 89.3%; User's = 87.7%).

@@

```
The comparisons of digitized versus segmented field boundaries showed that the mean field size across all validation sites averaged `r round(segha, 2)` ha (Median = `r round(seghamed, 2)`; StDev = `r round(seghasd, 2)`), which was `r round((segha - valha) / valha * 100) / 100` times larger than the `r round(valha, 2)` ha (Median = `r round(valhamed, 2)`; StDev = `r round(valhasd, 2)`) mean area of labeller-digitized polygons. This discrepancy was primarily caused by results in four AOIs (2, 3, 7, and 15; Figure S11), where segments averaged between `r round(lrgst_seg$rng[1], 2)` and `r round(lrgst_seg$rng[2], 2)` ha, compared to `r round(lrgst_val$rng[1], 2)` - `r round(lrgst_val$rng[2], 2)` ha for the corresponding hand-digitized polygons. The number of segmented fields per validation site averaged `r segn` (median = `r segnmed`; StDev = `r segnsd`) compared to `r valn` (median = `r valnmed`; StDev = `r valnsd`) for digitized polygons (Figure S12).

@@

The vectorized map provides additional information on how the characteristics of croplands can vary geographically, ranging from narrow, strip-like fields in parts of AOI 15 (Figure \ref{fig:mainmap}'s lower right inset) to more densely packed, less distinctly shaped fields in AOI 5 (upper right inset in Figure \ref{fig:mainmap}). To explore how field characteristics varied geographically, we mapped the average field size and total number of fields within each 0.05 degree tile (Figure S13). These patterns generally correspond to those seen in the cropland density map (Figure \ref{fig:mainmap}), with larger sizes and field counts generally occurring in areas of higher field density, although the biases inherent in both measures (Figures S11-12) complicate the interpretation of those variations. However, we can use the estimated biases to develop adjusted estimates of field sizes and counts for each AOI, and for Ghana overall (Table \ref{tab:sizentab}). These adjusted estimates show that the typical field size in Ghana is 1.73 ha, ranging from 0.96 in AOI 4 to 2.82 ha in AOI 4, with fields in the forest zone AOIs (10, 11, 13, 14, 16) generally smaller than those in the northern half of the country (Table \ref{tab:sizentab}). The total number of fields was estimated to be 1,662,281 overall, or 205 fields per tile on average, ranging from 108/tile in AOI 4 to 399/tile in AOI 6.

@@

In addition to the more detailed update of cropland extent, our maps also provide new information on the size and number of fields in Ghana (Figures \ref{fig:mainmap}, S11-12). Previous work to estimate such agricultural characteristics have often focused on farm, rather than field, size using census data [@jayneAfricaChangingFarm2016; @LowderNumberSizeDistribution2016; @SambergSubnationaldistributionaverage2016; @VonBraunSmallscalefarmersliberalized2004]. Efforts to map field boundaries in smallholder-dominated agricultural systems have either used *in situ* data collection [@CarlettoFactartifactimpact2013; @CarlettoGuesstimatesGPStimatesLand2015] or remote sensing studies over relatively small [e.g. @forkuorIntegrationOpticalSynthetic2014; @perselloDelineationAgriculturalFields2019] or discontiguous [@Estesplatformcrowdsourcingcreation2016] areas. The most extensive studies to date used crowdsourced volunteers to classify fields into broad size classes, based on their interpretations of imagery sampled from high resolution virtual globes [@lesivEstimatingGlobalDistribution2019; @FritzMappingglobalcropland2015]. Those efforts included country-specific results for Ghana (n = 263), which can be converted into an average field size estimate of `r round(lesiv_est, 2)` ha\footnote{Obtained by calculating the weighted mean from the count of the five size classes and the mean of the hectare range provided for the four smallest size classes, and the lower bound of the size range provided the largest size class. Data sourced from Table S3 in Lesiv et al. 2019.}. This estimate exceeds our Ghana-wide average segment size (3.92 ha; Table \ref{tab:sizentab}), but is closer to the mean (`r aoi_nofor` ha) within AOIs 1-9, 12, and 15, which is where most of the crowdsourced sample appears to have been collected. However, our bias-corrected estimates of 1.73 (Ghana-wide) and `r aoi_nofor_adj` (AOIs 1-9, 12, and 15) ha were much smaller.

@@

The patterns of cropland-class accuracies varied by zone. These zones largely align, albeit with some discrepancies, with the country's agroecozones, thus the accuracy patterns may be partially attributed to some regions being harder to map than others. Producer's accuracies for both maps were highest in the two northern zones (1 and 2), which are primarily savannas (Figure S10), and lowest in zones 3 and 4, which are comprised of forest or coastal savannas. User's accuracies followed a similar pattern, with the exception of Zone 3, which had the highest User's accuracy, albeit from a very small sample. Aligning the reference samples more precisely with agroecozone boundaries (Figure S10B) provides further insight into error patterns (Table S4). Coastal savannas in the southeast had the highest Producer's accuracy but lowest User's accuracy for the per-pixel map, presumably because this region's numerous areas of high density cropland, combined with low woody cover in surrounding uncultivated areas, helped to promote commission error. Maps in the two northern savanna agroecozones had the best balance between omission and commission error, and had the highest overall User's accuracy. The transitional (between forest and savanna) agroecozone had a very low Producer's accuracy (21%), which likely reflects the fact that it was divided between several AOIs for mapping (Figure S4), within which it typically covered a smaller share of area relative to the other agroecozones. This likely caused insufficient representation of this AEZ in training samples, particularly in AOIs 10 and 11 (Figure S4B).

@@

The inherent difficulty of the labelling task was another major limiting factor. Our system was designed to minimize the error inherent in labelling, but determining croplands from non-croplands in these agricultural systems can be a difficult task. Labellers have to evaluate multiple image sources and to rely heavily on judgement, which inevitably leads to errors. Interpretation is particularly hard where the background savanna vegetation and croplands have similar reflectance during the dry season, which is a particular problem in AOIs 2 and 3. Smaller field sizes also complicate labelling, as these become increasingly indistinct in the ~4 m PlanetScope composites. The difficulty of labelling is reflected in the magnitude of the Bayesian Risk metrics (Figure S6), and by the average score achieved by each labeller against our training reference dataset (71%; Figure S5B). Although prior work [@mellorExploringIssuesTraining2015; @rodriguez-galianoAssessmentEffectivenessRandom2012] found that Random Forests are robust to label errors, we found that they have substantial impact (Figure \ref{fig:trainingimpact}), which suggest that simply improving label quality may be one of the single most important investments towards improved model accuracy.

@@

## Error mitigation features
Despite the error sources mentioned above, several features of the system proved effective in mitigating error, leading to a higher overall accuracy than would have otherwise been possible. Label accuracy assessment and consensus labelling appeared to be the most effective error mitigation tools. Label accuracy measures allowed us to quantify the substantial impact of label error on model performance (Figure \ref{fig:trainingimpact}), while consensus labels substantially reduced individual labelling errors, resulting in maps that were more accurate than they would have been had we used individually generated labels. Labeller-specific accuracy measures also helped to improve the quality of the consensus labels, by placing higher weight on labels more likely to be accurate during the merging process, rather than giving equal weight to potentially less accurate labels. The ability to select the most accurate individual labels for a site also allowed us to develop independent estimates of field size to which measures of confidence can be attached (Figure S5B), which we were in turn able to correct estimates of field sizes and numbers (Table \ref{tab:sizentab}).
